---
title: "SwiFCoIBMove: Data import and processing"
author: "Cédric Scherer"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    link-citations: yes
editor_options: 
  chunk_output_type: console
---
  
```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


Script to import, process, and subset the simulation results (saved as .csv by running simulations using the "SwiFCoIBM-move.nlogo" file in the model folder).


# Setup

```{r libraries, message = F}
## libraries
library(tidyverse)     ## data wrangling
library(magrittr)      ## ceci n'est pas une pipe

filedate <- "2019-03-05"
```


# Data

## Load data

```{r load-data}
df_all <- df_raw <- data.table::fread("./simulations/2019-03-05_SwiFCoIBMove_mue4.csv", skip = 6)
```

## Data cosmetics

### rename columns

```{r rename-cols}
rename_map <- c(
  '[run number]'                                                                      = "run",
  'case_fatality'                                                                     = "cfr",
  'file'                                                                              = "scenario",
  'roaming'                                                                           = "roaming",
  'mue'                                                                               = "mue",
  'mue_max'                                                                           = "mue_max_rm",
  'q'                                                                                 = "dir_pers",
  'run_years'                                                                         = "run_years_rm",
  'seed_setup'                                                                        = "seed_setup_rm",   
  'mean_quality'                                                                      = "quality_rm",
  'herd_prop'                                                                         = "herd_prop_rm",
  'release_fct'                                                                       = "release_fct_rm",
  'longevity'                                                                         = "longevity_rm",
  'age_blur'                                                                          = "age_blur_rm",
  'fem_prob'                                                                          = "fem_prob_rm",
  'dist_disp'                                                                         = "dist_disp_rm",
  'year_release'                                                                      = "year_release_rm",
  'fert_red'                                                                          = "fert_red_rm",
  'fetal_inf'                                                                         = "fetal_inf_rm",
  't_anti'                                                                            = "t_anti_rm",
  't_trans'                                                                           = "t_trans_rm",
  'b_within'                                                                          = "beta_within",
  'b_between'                                                                         = "beta_between_rm",
  'b_move'                                                                            = "beta_move",
  '[step]'                                                                            = "week",
  'seed'                                                                              = "seed",
  'count turtles'                                                                     = "ind_all",
  'count turtles with [epi_stat = \"\"esSusc\"\"]'                                    = "ind_susc",
  'count turtles with [epi_stat = \"\"esTrans\"\"]'                                   = "ind_trans",
  'count turtles with [epi_stat = \"\"esLeth\"\"]'                                    = "ind_leth",
  'count turtles with [dem_stat = \"\"dsRoaming\"\"]'                                 = "roam_all",
  'count turtles with [dem_stat = \"\"dsRoaming\"\" AND epi_stat = \"\"esSusc\"\"]'   = "roam_susc",
  'count turtles with [dem_stat = \"\"dsRoaming\"\" AND epi_stat = \"\"esTrans\"\"]'  = "roam_trans",
  'count turtles with [dem_stat = \"\"dsRoaming\"\" AND epi_stat = \"\"esLeth\"\"]'   = "roam_leth",
  'new_trans'                                                                         = "new_trans",
  'new_leth'                                                                          = "new_leth",
  'contacts_avg'                                                                      = "contacts_avg",
  'contacts_med'                                                                      = "contacts_med",
  'contacts_lwr'                                                                      = "contacts_lwr",
  'contacts_upr'                                                                      = "contacts_upr",
  'contacts_max'                                                                      = "contacts_max",
  'contacts_var'                                                                      = "contacts_var",
  'contacts_1'                                                                        = "contacts_1",
  'contacts_2'                                                                        = "contacts_2",
  'contacts_3'                                                                        = "contacts_3",
  'contacts_4'                                                                        = "contacts_4",
  'contacts_5'                                                                        = "contacts_5",
  'contacts_6'                                                                        = "contacts_6",
  'contacts_7'                                                                        = "contacts_7",
  'contacts_8'                                                                        = "contacts_8",
  'contacts_9'                                                                        = "contacts_9",
  'trans_w'                                                                           = "trans_w",
  'trans_b'                                                                           = "trans_b",
  'trans_g'                                                                           = "trans_g",
  'patches_0'                                                                         = "patches_0",
  'patches_1'                                                                         = "patches_1",
  'patches_2'                                                                         = "patches_2",
  'patches_3'                                                                         = "patches_3",
  'patches_4'                                                                         = "patches_4",
  'patches_5'                                                                         = "patches_5",
  'patches_6'                                                                         = "patches_6",
  'patches_7'                                                                         = "patches_7",
  'patches_8'                                                                         = "patches_8",
  'patches_9'                                                                         = "patches_9",
  'count patches with [is_infected = 1]'                                              = "cell_infected",
  'count patches with [is_infectious = 1]'                                            = "cell_infectious",
  'max [id] of patches'                                                               = "n_cluster",
  'count_init'                                                                        = "count_init",
  'count_init_roaming'                                                                = "count_init_roaming",
  'dens_var'                                                                          = "dens_var_all",
  'dens_roam_var'                                                                     = "dens_var_roam",
  'dens_inf_group_var'                                                                = "dens_var_inf_group",
  'dens_inf_roam_var'                                                                 = "dens_var_inf_roam",
  'quality_mean_6'                                                                    = "qual_mean_6",
  'mean [visited] of turtles with [dem_stat = \"\"dsRoaming\"\"]'                     = "visits_mean",
  'F_infectious'                                                                      = "F_infectious",
  'F_infected'                                                                        = "F_infectected",
  'dist_inf'                                                                          = "dist_inf",
  'week_inf_max'                                                                      = "week_inf_max",
  'week_release'                                                                      = "week_release",
  'week_last'                                                                         = "week_last"
)
names(df_all) <- rename_map[names(df_all)]
rm(rename_map)
```

### select columns used for further analyses and factorize variables

```{r select+factorize}
df_all %<>% 
  dplyr::select(-ends_with("_rm")) %>% 
  mutate_at(c(1:2, 5:length(.)), as.numeric)
 
(lev_roam <- unique(df_all$roaming))
(lev_scen <- unique(df_all$scenario))

df_all %<>% 
  mutate(
    run      = factor(run),
    cfr      = factor(cfr),
    roaming  = factor(roaming, levels = lev_roam),
    scenario = factor(scenario, levels = lev_scen)
  ) 
```

### correct non-logical values, calculate additional parameters and fill up for each measurement if needed

```{r correct+add}
df_all %<>% 
  group_by(run) %>% 
  filter(week != 0) %>% 
  mutate(
    ind_immu     = ind_all - ind_trans - ind_leth - ind_susc,
    ind_inf      = ind_trans + ind_leth,
    roam_inf     = roam_trans + roam_leth,
    roam_immu    = roam_all - roam_trans - roam_leth - roam_susc,
    new_inf      = new_trans + new_leth,
    week_last    = max(week_last),
    trans_m      = trans_w + trans_b,
    trans        = trans_g + trans_m,
    contacts_avg = na_if(contacts_avg, -999),
    contacts_med = na_if(contacts_med, -999),
    contacts_lwr = na_if(contacts_lwr, -999),
    contacts_upr = na_if(contacts_upr, -999),
    contacts_max = na_if(contacts_max, -999),
    contacts_var = na_if(contacts_var, -999)
  ) %>%
  filter(week <= week_last + 1) %>% 
  mutate(
    week_ext = if_else(week_last == 624, NA_real_, week_last + 1),
    week_pathogen = week - week_release,
    week_inf_max = week_inf_max - week_release + 1, 
    week_inf_max = ifelse(week_inf_max < 1, NA, week_inf_max),
    duration_weeks = max(week_pathogen),
    duration_quarter = duration_weeks / 13
  ) %>% 
  arrange(run, week) %>%
  tbl_df()
```

## Save as .Rds
 
```{r save-proc }
saveRDS(df_all, file = glue::glue("./data/{filedate}_SwiFCoIBMove_all.Rds"))
```


# Subsets

## Create subsets

```{r create-subsets}
## only outbreak weeks
df_inf <- df_all %>% 
  filter(week >= week_release)

## mean values for time of outbreak
df_infmean <- df_inf %>% 
  group_by(run) %>%
  dplyr::select(-week, -week_pathogen, -week_ext) %>% 
  summarise_all(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else first(.)))

## sample data for testing
df_sample <- df_all %>%
  group_by(roaming, scenario, mue, cfr, dir_pers) %>%
  do(sample_n(., 500))
```

## Save subsets as .Rds

```{r save-subsets}
saveRDS(df_inf,     file = glue::glue("./data/{filedate}_SwiFCoIBMove_inf.Rds"))
saveRDS(df_infmean, file = glue::glue("./data/{filedate}_SwiFCoIBMove_infmean.Rds"))
```


# Data info

## Check sample size

```{r sample-size}
df_n <- df_infmean %>%   
  group_by(roaming, cfr, mue, scenario) %>%
  summarise(n = n())

length(unique(df_n$n))
```

## Variables

```{r variables}
## number of unique values
df_infmean %>% 
  keep(is.factor) %>% 
  map_int(n_distinct) 

## unique values of each variable
df_infmean %>% 
  keep(is.factor) %>% 
  dplyr::select(-run) %>% 
  map(unique)
```

***

# Version info

```{r version}
version
sessionInfo()
```
