---
title: "SwiFCoIBMove: Figures"
author: "Cédric Scherer"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    link-citations: yes
editor_options: 
  chunk_output_type: console
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Preparation

## Setup

```{r setup}
## libraries
library(tidyverse)  ## data wrangling
library(magrittr)   ## ceci n'est pas une pipe
library(glue)       ## glue strings together
library(patchwork)  ## glue ggplots together
library(viridis)    ## viridis color palettes
library(ggridges)   ## how to draw a circle: geom_ridgeline_gradient
library(EnvStats)   ## add sample size to ggplots
library(ggbeeswarm) ## make beeswarm plots

## ggplot theme
source("./R/ggtheme_grey.R")

## enable sourcing of .Rmd files
source("./R/source-rmd.R")
```


## Data

```{r data}
filedate <- "2019-03-05"

## data
destfile1 = glue::glue("./data/{filedate}_SwiFCoIBMove_inf.Rds")
destfile2 = glue::glue("./data/{filedate}_SwiFCoIBMove_infmean.Rds")
destfile3 = glue::glue("./data/{filedate}_SwiFCoIBMove_all.Rds")

if (!file.exists(destfile1) | 
    !file.exists(destfile2) | 
    !file.exists(destfile3)) {
  print("reading and cleaning raw data")
  source_rmd("2_data.Rmd")
} else if (exists("df_inf") && 
           exists("df_infmean") && 
           exists("df_all")) {
  print("data available")
} else {
  print("loading data")
  df_inf     <- readRDS(destfile1)
  df_infmean <- readRDS(destfile2) 
  df_all     <- readRDS(destfile3)
}
```


## Prepare data for plotting

```{r data-preparation}
## set clean labels for factors
lab_roam <- c("Correlated random walk", "Habitat-dependent movement", "Competition-driven movement")
lab_roam_n <- c("Correlated\nrandom walk", "Habitat-dependent\nmovement", "Competition-driven\nmovement")
lab_scen <- c("Homogeneous", "Large clusters", "Small clusters", "Random") 

df_inf_plots <- df_inf %>% 
  mutate(
    roaming = factor(roaming, labels = lab_roam),
    scenario = factor(scenario, levels = c("hom", "patch_l", "patch_s", "rand"), labels = lab_scen),
    week_ext = if_else(is.na(week_ext), 625, week_ext)
  ) %>% 
  ungroup()

df_infmean_plots <- df_infmean %>%
  mutate(
    roaming = factor(roaming, labels = lab_roam),
    scenario = factor(scenario, levels = c("hom", "patch_l", "patch_s", "rand"), labels = lab_scen)
  ) %>% 
  group_by(roaming, scenario) %>% 
  mutate(
    pers = if_else(duration_weeks > 10*52, 1, 0),
    inv = if_else(duration_weeks > 2.5*52, 1, 0)) %>% 
  ungroup()

## labels for weeks and years
labs_weeks <- tibble("week" = seq(0, 572, by = 52)) %>% 
  mutate(
    weeks = ifelse(week %% 104 == 0, glue::glue("{week}"), ""),
    weeks_year = ifelse(week %% 104 == 0, glue::glue("{week}\nyear {week/52}"), "")
  ) %>% 
  filter(week <= 520)

weeks_year <- labs_weeks %>% pull(weeks_year)
weeks_only <- labs_weeks %>% pull(weeks)
```


## Variables & Settings

Plots visualize results for a case fatality risk of 0.5 (1:1 chance to become either transient or lethally infected). 

```{r plot-settings}
## case fatality risk to visualize
cfr_filter <- 0.5

## scenario to visualize (i.e. Fig. 2 single and Fig. 3)
scenario_filter <- "Small clusters"

## save plots?
save <- TRUE

## quality of plots
dpi <- 750
```


# Figures Main Text

## Fig. 1: Scheme

### a) General dynamics

```{r fig-1a-pop-dynamics-single, fig.width = 3.6, fig.height = 2.5}
fig_1a <- df_all %>%
  filter(
    cfr == cfr_filter,
    scenario == "patch_s",
    roaming == "CD",
    duration_weeks > 520
  ) %>% 
  filter(run == 2008) %>% 
  dplyr::select(
    week = week,
    S = ind_susc,
    I = ind_inf,
    R = ind_immu
  ) %>% 
  gather(class, count, -week) %>% 
  mutate(
    class = factor(class, levels = c("S", "R", "I")),
    count = ifelse((class %in% c("R", "I") & week < 95), NA_real_, count)  ## 95 == week_release
  ) %>% 
  ggplot(aes(week, count)) +   
    geom_rect(aes(xmin = 53, xmax = 104, ymin = -Inf, ymax = Inf), fill = "grey90") +
    geom_hline(yintercept = 0, color = "grey85", size = 0.3) +
    geom_line(aes(color = class), size = 0.5) +
    geom_vline(xintercept = 95, linetype = "dotted", color = "grey50", size = 0.4) +  ## 95 == week_release
    scale_x_continuous(breaks = seq(0, 624, by = 52), limits = c(0, 624), 
                       labels = c("0", "", "2", "", "4", "", "6", "", "8", "", "10", "", "12")) + 
    scale_y_continuous(labels = scales::comma_format(),
                       expand = c(0.05, 0.05),
                       limits = c(0, 38000)) +
    scale_color_manual(name = NULL, values = c("grey20", "navajowhite3", "orangered")) +
    theme(legend.position = c(0.6, 0.93), 
          legend.box = "horizontal",
          legend.background = element_rect(fill = "transparent"),
          legend.text = element_text(face = "bold"),
          plot.background = element_rect(fill = "transparent", colour = NA)) +
    labs(x = "Simulated years", y = "Individuals", title = NULL) +
    guides(color = guide_legend(nrow = 1))
```

### b) Infections over time

```{r fig-1b-inf-dynamics-single, fig.width = 3.7, fig.height = 2.5}
df_1b <- df_inf_plots %>% 
  filter(
    cfr == cfr_filter,
    scenario == "Small clusters",
    roaming == "Competition-driven movement"
  ) %>% 
  mutate(
    pers = ifelse(duration_weeks > 520, ">10", "<10")
  ) %>% 
  group_by(scenario, roaming, week_pathogen) %>% 
  mutate(
    meanInf  = mean(ind_inf),
    duration_weeks = if_else(duration_weeks == week_pathogen, duration_weeks, NA_real_)
  )

labs <- c("\u2264 10 years", "> 10 years")

fig_1b <- ggplot() +   
  geom_hline(yintercept = 0, color = "grey85", size = 0.3) +
  geom_line(data = filter(df_1b, pers == ">10"), 
            aes(week_pathogen, ind_inf, 
                group = run, color = ">10"), 
            size = 0.33, alpha = 0.4) +
  geom_line(data = filter(df_1b, pers == "<10"), 
            aes(week_pathogen, ind_inf, 
                group = run, color = "<10"), 
            size = 0.33, alpha = 0.4) +
  geom_line(data = df_1b, 
            aes(week_pathogen, meanInf), 
            color = "black", size = 0.35) +
  geom_rug(data = filter(df_1b, pers == "<10"), 
           aes(duration_weeks), 
           sides = "b", alpha = 0.4, size = 0.3) +
  scale_x_continuous(breaks = seq(0, 520, by = 52), 
                     limits = c(0, 520), 
                     labels = 0:10) + 
  scale_y_continuous(labels = scales::comma_format(), 
                     expand = c(0.05, 0.05)) +
  scale_color_manual(values = c("<10" = "tan1", ">10" = "grey20"), 
                     name = NULL, labels = labs) +
  guides(color = guide_legend(override.aes = list(size = 0.8, alpha = 1))) +
  theme(legend.position = c(0.65, 0.85), 
        legend.box = "horizontal",
        legend.background = element_rect(fill = "transparent"),
        legend.key.width = unit(1.8, "lines"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(x = "Years since pathogen release", y = "Infected individuals", title = NULL)
```

### c) Outbreak duration

```{r fig-1c-duration-single, fig.width = 3.5, fig.height = 2.5}
n <- df_infmean_plots %>%   
  group_by(roaming, scenario, cfr) %>%
  count() %>% 
  ungroup() %>% 
  summarize(n = unique(n)) %>% 
  pull()

df_1c <- df_infmean_plots %>% 
  filter(
    cfr == cfr_filter,
    scenario == "Small clusters",
    roaming == "Competition-driven movement"
  ) %>%
  mutate(duration_quarter = if_else(duration_quarter > 40, 48, floor(duration_quarter))) %>% 
  group_by(roaming, cfr, scenario, duration_quarter) %>% 
  summarize(prop = n() / n)

fig_1c <- ggplot(filter(df_1c, duration_quarter != 48), aes(duration_quarter / 4, prop)) + 
  geom_hline(yintercept = 0, color = "grey85", size = 0.3) +
  geom_bar(aes(group = duration_quarter), fill = "grey30", stat = "identity", width = 0.2) +
  geom_bar(data = filter(df_1c, duration_quarter == 48), 
           aes(group = duration_quarter), fill = "tan1", stat = "identity", width = 0.2) +
  geom_vline(xintercept = 11, linetype = "dotted", size = 0.5) +
  geom_vline(xintercept = c(2.625, 5.125, 7.625), linetype = "dotted", color = "grey30", size = 0.4) +
  scale_x_continuous(breaks = c(0, 2.5, 5, 7.5, 10, 12), 
                     labels = c(0, 2.5, 5, 7.5, 10, ">10")) +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3), expand = c(0.01, 0.01)) +
  expand_limits(x = c(0, 12.25), y = c(0, 0.22)) +
  theme(legend.position = "none",
        plot.background = element_rect(fill = "transparent",colour = NA)) +
  labs(x = "Outbreak duration (years)", y = "Proportion of runs", title = NULL)
```

### Full panel

```{r fig-1-full-panel, fig.width = 11.7, fig.height = 2.5}
((fig_1a + plot_spacer() + fig_1b + plot_spacer() + fig_1c) * 
    theme(line = element_line(color = "black"), 
          rect = element_rect(color = "black"), 
          panel.border = element_rect(color = "black")) + 
   plot_layout(widths = c(1.3, 0.02, 1.4, 0.02, 1.5), nrow = 1))

if (save == T) { 
  ggsave("./plots/final/Fig1_workflow.png", width = 11.7, height = 3, dpi = dpi)
  ggsave("./plots/final/Fig1_workflow.pdf", width = 11.7, height = 3, device = cairo_pdf) 
}
```


## Fig. 3: Persistence probabilities

```{r fig-3-pers-groups, fig.width = 13.5, fig.height = 7.5}
df_3 <- df_infmean_plots %>% 
  filter(cfr == cfr_filter)

## per movement rule
fig_3_roam <- 
  df_3 %>% 
  group_by(roaming) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 1.96),
    pers_upp = pers_mean + (pers_se * 1.96),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 1.96),
    inv_upp = inv_mean + (inv_se * 1.96)
  ) %>%
  mutate(group = as.factor("All landscape scenarios")) %>%
  ungroup() %>% 
    ggplot(aes(x = roaming, color = roaming)) +   
    geom_bar(aes(y = pers_mean, fill = roaming), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.05, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp),  width = 0.1, size = 0.25, show.legend = F) +
    facet_wrap(~ group) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "right",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_manual(values = c("grey70", "grey45", "grey20"), name = "Movement rule") + 
    scale_fill_manual(values = c("grey80", "grey55", "grey30"), name = "Movement rule") + 
    labs(x = "Movement rule", y = "Persistence probability", tag = "(a)  ")

## per landscape scenario
fig_3_scen <- 
  df_3 %>% 
  group_by(scenario) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 1.96),
    pers_upp = pers_mean + (pers_se * 1.96),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 1.96),
    inv_upp = inv_mean + (inv_se * 1.96)
  ) %>%
  mutate(group = as.factor("All movement\nrules")) %>% 
  ungroup() %>% 
  ggplot(aes(x = scenario, color = scenario)) +   
    geom_bar(aes(y = pers_mean, fill = scenario), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.15, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp),  width = 0.3, size = 0.25, show.legend = F) +
    facet_wrap(~ group) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_viridis_d(option = "inferno", begin = 0.05, end = 0.75, direction = -1, guide = F) +
    scale_fill_viridis_d(option = "inferno", begin = 0.15, end = 0.85, direction = -1, guide = F) +
    labs(x = "Landscape scenario", y = "Persistence probability", tag = "(b)  ")

fig_3_detail <- df_3 %>% 
  group_by(roaming, scenario) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 1.96),
    pers_upp = pers_mean + (pers_se * 1.96),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 1.96),
    inv_upp = inv_mean + (inv_se * 1.96)
  ) %>%
  ungroup() %>% 
  mutate(roaming = factor(roaming, labels = lab_roam_n)) %>% 
  ggplot(aes(x = scenario, color = scenario)) + 
    geom_bar(aes(y = pers_mean, fill = scenario), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.15, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp), width = 0.3, size = 0.25, show.legend = F) +
    facet_wrap(~ roaming) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "right",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_viridis_d(option = "inferno", begin = 0.05, end = 0.75, direction = -1, name = "Landscape scenario") +
    scale_fill_viridis_d(option = "inferno", begin = 0.15, end = 0.85, direction = -1, name = "Landscape scenario") +
    labs(x = "Landscape scenario", y = "Persistence probability", tag = "(c)  ")

(fig_3 <- plot_spacer() + fig_3_roam + fig_3_scen + fig_3_detail + 
            plot_layout(nrow = 2, heights = c(1, 1.5), widths = c(1, 3)))

if (save == T) { 
  ggsave("./plots/final/Fig3_pers_groups_95.png", width = 13.5, height = 7.5, dpi = dpi) 
  ggsave("./plots/final/Fig3_pers_groups_95.pdf", width = 13.5, height = 7.5, device = cairo_pdf)
}
```

```{r fig-3-pers-groups-CI50, fig.width = 13.5, fig.height = 7.5}
## per movement rule
fig_3_roam_50 <- 
  df_3 %>% 
  group_by(roaming) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 0.675),
    pers_upp = pers_mean + (pers_se * 0.675),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 0.675),
    inv_upp = inv_mean + (inv_se * 0.675)
  ) %>%
  mutate(group = as.factor("All landscape scenarios")) %>%
  ungroup() %>% 
    ggplot(aes(x = roaming, color = roaming)) +   
    geom_bar(aes(y = pers_mean, fill = roaming), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.05, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp),  width = 0.1, size = 0.25, show.legend = F) +
    facet_wrap(~ group) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "right",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_manual(values = c("grey70", "grey45", "grey20"), name = "Movement rule") + 
    scale_fill_manual(values = c("grey80", "grey55", "grey30"), name = "Movement rule") + 
    labs(x = "Movement rule", y = "Persistence probability", tag = "(a)  ")

## per landscape scenario
fig_3_scen_50 <- 
  df_3 %>% 
  group_by(scenario) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 0.675),
    pers_upp = pers_mean + (pers_se * 0.675),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 0.675),
    inv_upp = inv_mean + (inv_se * 0.675)
  ) %>%
  mutate(group = as.factor("All movement\nrules")) %>% 
  ungroup() %>% 
  ggplot(aes(x = scenario, color = scenario)) +   
    geom_bar(aes(y = pers_mean, fill = scenario), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.15, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp),  width = 0.3, size = 0.25, show.legend = F) +
    facet_wrap(~ group) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_viridis_d(option = "inferno", begin = 0.05, end = 0.75, direction = -1, guide = F) +
    scale_fill_viridis_d(option = "inferno", begin = 0.15, end = 0.85, direction = -1, guide = F) +
    labs(x = "Landscape scenario", y = "Persistence probability", tag = "(b)  ")

fig_3_detail_50 <- df_3 %>% 
  group_by(roaming, scenario) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv),
    all_n = n(),
    ## CIs for persistence
    pers_n = sum(pers),
    pers_se = (sqrt(pers_mean * (1 - pers_mean) / all_n)),
    pers_low = pers_mean - (pers_se * 0.675),
    pers_upp = pers_mean + (pers_se * 0.675),
    ## CIs for invasion
    inv_n = sum(inv),
    inv_se = (sqrt(inv_mean * (1 - inv_mean) / all_n)),
    inv_low = inv_mean - (inv_se * 0.675),
    inv_upp = inv_mean + (inv_se * 0.675)
  ) %>%
  ungroup() %>% 
  mutate(roaming = factor(roaming, labels = lab_roam_n)) %>% 
  ggplot(aes(x = scenario, color = scenario)) + 
    geom_bar(aes(y = pers_mean, fill = scenario), stat = "identity", position = "dodge", width = 0.7, color = NA) + 
    geom_bar(aes(y = inv_mean), stat = "identity", position = "dodge", width = 0.7, fill = "transparent", size = 0.25) + 
    geom_errorbar(aes(ymin = inv_low, ymax = inv_upp), width = 0.15, size = 0.25, show.legend = F) +
    geom_errorbar(aes(ymin = pers_low, ymax = pers_upp), width = 0.3, size = 0.25, show.legend = F) +
    facet_wrap(~ roaming) +
    expand_limits(y = c(0, 1)) +
    theme(legend.position = "right",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_text(size = 14),
          axis.title.y = element_text(size = 14)) +
    scale_color_viridis_d(option = "inferno", begin = 0.05, end = 0.75, direction = -1, name = "Landscape scenario") +
    scale_fill_viridis_d(option = "inferno", begin = 0.15, end = 0.85, direction = -1, name = "Landscape scenario") +
    labs(x = "Landscape scenario", y = "Persistence probability", tag = "(c)  ")

(fig_3_50 <- plot_spacer() + fig_3_roam_50 + fig_3_scen_50 + fig_3_detail_50 + 
               plot_layout(nrow = 2, heights = c(1, 1.5), widths = c(1, 3)))

if (save == T) { 
  ggsave("./plots/final/Fig3_pers_groups_50.png", width = 13.5, height = 7.5, dpi = dpi) 
  ggsave("./plots/final/Fig3_pers_groups_50.pdf", width = 13.5, height = 7.5, device = cairo_pdf)
}
```

## Fig. 4: Temporal outbreak dynamics

```{r fig-4-temp-dynamics, fig.width = 8.9, fig.height = 6.9}
## normed transmission per habitat quality - over time
df_4 <- df_inf_plots %>% 
  filter(
    cfr == cfr_filter,
    week_pathogen > 0, 
    week_pathogen < 521
  )

df_trans <- df_4 %>%
  filter(duration_weeks > 260) %>% 
  dplyr::select(
    run,
    week_pathogen,
    roam_all,
    scenario,
    roaming,
    trans_g,
    trans_m,
    trans_w,
    trans_b
  ) %>% 
  gather(trans, trans_count, -c(run:roaming)) %>% 
  mutate(
    trans = substr(trans, nchar(trans), nchar(trans) + 1),
    trans_count = trans_count / roam_all,
    quarter = week_pathogen %/% 13
  ) %>% 
  filter(trans > 0) %>% 
  group_by(roaming, scenario, quarter, trans) %>%
  summarize(
    avg = mean(trans_count, na.rm = T), 
    upr = avg + sd(trans_count, na.rm = T), 
    lwr = avg - sd(trans_count, na.rm = T)
  ) %>%
  ungroup() %>% 
  mutate(
    lwr = ifelse(lwr < 0, 0, lwr),
    avg = ifelse(trans == "b" & scenario == "Homogeneous", NA, avg),
    lwr = ifelse(trans == "b" & scenario == "Homogeneous", NA, lwr),
    upr = ifelse(trans == "b" & scenario == "Homogeneous", NA, upr),
    type = case_when(trans == "g" ~ "Group-living hosts",
                     trans == "m" ~ "Ranging hosts",
                     trans == "w" ~ "Within-cluster",
                     trans == "b" ~ "Between-cluster")
  ) 
  
n <- df_infmean_plots %>% 
  filter(cfr == cfr_filter) %>%   
  group_by(roaming, scenario) %>%
  count() %>% 
  ungroup() %>% 
  summarize(n = unique(n)) %>% 
  pull()

df_dur <- df_infmean_plots %>% 
  filter(
    cfr == cfr_filter,
    duration_weeks < 521
  ) %>% 
  mutate(duration_quarter = if_else(duration_quarter > 40, 50, floor(duration_quarter))) %>% 
  group_by(roaming, scenario, duration_quarter) %>% 
  summarize(prop = n() / n)

df_ext <- df_4 %>% 
  group_by(roaming, scenario, run) %>% 
  summarize(week_ext = max(week_ext - week_release)) %>% 
  filter(week_ext < 520) %>% 
  mutate(avg = 0)

tm <- df_infmean_plots %>% 
  filter(
    cfr == cfr_filter,
    duration_weeks <= 520
  ) %>%   
  group_by(roaming, scenario) %>%
  summarize(tm = mean(duration_weeks, na.rm = T) / 52)

(fig_4 <- df_trans %>% 
  filter(trans %in% c("w", "b")) %>%
  ggplot(aes(quarter / 4, avg)) +
    geom_bar(data = filter(df_dur, duration_quarter < 50), aes(duration_quarter / 4 + 0.125, prop / 5), 
             stat = "identity", width = 0.2, fill = "grey85") +
    geom_vline(data = tm, aes(xintercept = tm), linetype = "dashed", color = "grey60", size = 0.3) +
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = type), alpha = 0.25) + 
    geom_line(aes(color = type), size = 0.5) +
    geom_rug(data = df_ext, aes(x = week_ext / 52), 
           sides = "b", color = "grey60", alpha = 0.4, size = 0.4) +
    facet_grid(scenario ~ roaming) +
    scale_x_continuous(breaks = seq(0, 10, by = 2.5), labels = seq(0, 10, by = 2.5)) +
    scale_y_continuous(limits = c(-0.001, 0.055), breaks = seq(0, 0.06, by = 0.02), 
                       sec.axis = sec_axis(~ . * 5)) +
    scale_color_manual(values = c("#f6b233", "#6a009a"), name = NULL) +
    scale_fill_manual(values  = c("#f5980a", "#a332d7"), name = NULL) +
    labs(x = "Outbreak duration (years)", 
         y = "Weekly transmissions per moving host") +
    theme(legend.position = "bottom",
          strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          strip.text = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          axis.text = element_text(size = 10),
          legend.text = element_text(size = 10.5)))

if (save == T) {  
  ggsave("./plots/final/Fig4_dynamics_b+w.png", width = 8.9, height = 6.9, dpi = dpi)
  ggsave("./plots/final/Fig4_dynamics_b+w.pdf", width = 8.9, height = 6.9, device = cairo_pdf)
}
```


## Fig. 5: Multipanel movement + landscape consequences

```{r fig-5-MultiBox-Data}
## data set quarterly
df_quarter <- df_inf_plots %>% 
  filter(
    cfr == cfr_filter
  ) %>% 
  mutate(
    roaming = factor(roaming, labels = lab_roam_n),
    quarter = week %/% 13 + 1
  ) %>% 
  dplyr::select(
    run, roaming, scenario, ## vars
    week_pathogen, quarter, duration_weeks, ## time
    dens_var_all, dens_var_roam,  ## a: density of hosts
    visits_mean,  ## b: visits by ranging males
    contacts_1:contacts_9, patches_1:patches_9, roam_inf, ## c: quality-dependent contact rates
    cell_infectious, patches_0,  ## d: infected patches
    cell_infected, ## e: affected patches
    trans_w, trans_b, ind_inf, ind_all ## f: within- and between-patch trans
  ) %>% 
  mutate(
    dens_var_all = ifelse(dens_var_all == 0, NA, dens_var_all),
    dens_var_roam = ifelse(dens_var_roam == 0, NA, dens_var_roam)
  )

## Data for a) density of hosts
df_m1 <- df_quarter %>% 
  group_by(roaming, scenario, run, quarter) %>% 
  summarize(
    all = median(dens_var_all) / 4,
    males = median(dens_var_roam)
  ) %>% 
  gather(type, avg, -roaming, -scenario, -run, -quarter) %>% 
  mutate(type = factor(type, levels = c("all", "males"), 
                       labels = c("All hosts", "Moving hosts")))

## Data for b) visits by ranging males
df_m2 <- df_all %>% 
  filter(cfr == 0.5) %>% 
  mutate(
    type = ifelse(week < week_release, "Pre-Outbreak", "Outbreak"),
    type = factor(type, levels = c("Pre-Outbreak", "Outbreak")),
    quarter = week %/% 13 + 1
  ) %>% 
  group_by(roaming, scenario, type, run, quarter) %>% 
  summarize(med = median(visits_mean, na.rm = T)) %>% 
  group_by(roaming, scenario, type)

## Data for c) contact rates per quality class
df_m3 <- df_quarter %>% 
  filter(roam_inf > 0) %>% 
  mutate(
    patches_low = patches_1 + patches_2 + patches_3,
    patches_med = patches_4 + patches_5 + patches_6,
    patches_hig = patches_7 + patches_8 + patches_9,
    contacts_low = (contacts_1 + contacts_2 + contacts_3) / (patches_low / 1250) / roam_inf,
    contacts_med = (contacts_4 + contacts_5 + contacts_6) / (patches_med / 1250) / roam_inf,
    contacts_hig = (contacts_7 + contacts_8 + contacts_9) / (patches_hig / 1250) / roam_inf
  ) %>% 
  group_by(roaming, scenario, run, quarter) %>% 
  summarize(
    low = median(contacts_low),
    med = median(contacts_med),
    hig = median(contacts_hig)
  ) %>% 
  gather(type, med, -roaming, -scenario, -run, -quarter) %>% 
  mutate(type = factor(type, levels = c("low", "med", "hig"), 
                             labels = c("Low ", "Medium", "High")))

hom <- df_m3  %>% 
  group_by(roaming, scenario) %>% 
  summarize(mean = mean(med, na.rm = T))

## Data for d) simultaneously affected patches
df_m4 <- df_quarter %>% 
  group_by(roaming, scenario, run, quarter) %>%  
  summarize(cell_infectious = mean(cell_infectious)) %>% 
  group_by(roaming, scenario, run) %>%  
  summarize(max = max(cell_infectious))

## Data for e) affected patches
df_m5 <- df_quarter %>% 
  group_by(roaming, scenario, run) %>%  ## no quarter since unique values per run
  summarize(max = max(cell_infected / (1250 - patches_0)))

## Data for f) trans within + between
df_m6 <- df_quarter %>%
  filter(week_pathogen > 52) %>% 
  dplyr::select(
    run, quarter, scenario, roaming, 
    trans_w, trans_b, patches_0
  ) %>% 
  mutate(
    trans_w = trans_w / (1250 - patches_0),
    trans_b = trans_b / (1250 - patches_0),
    trans_a = trans_w + trans_b
  ) %>% 
  group_by(roaming, scenario, run, quarter) %>% 
  summarize(
    trans_w = mean(trans_w, na.rm = T),
    trans_b = mean(trans_b, na.rm = T),
    trans_a = mean(trans_a, na.rm = T)
  ) %>% 
  mutate(trans_w = trans_w / trans_a, trans_b = trans_b / trans_a) %>% 
  dplyr::select(-trans_a) %>% 
  gather(type, avg, -roaming, -scenario, -run, -quarter) %>% 
  mutate(
    type = substr(type, nchar(type), nchar(type) + 1),
    avg = case_when(
      scenario == "Homogeneous" & type == "b" ~ 0,
      scenario == "Homogeneous" & type == "b" ~ 1,
      TRUE ~ avg
    )
  ) %>% 
  mutate(type = factor(type, levels = c("w", "b"),
                             labels = c("Within-cluster", "Between-cluster")))
```

### Fig. 5: Box + Beeswarm

```{r fig-5-MultiBox, fig.width = 9, fig.height = 11.5}
## a) density of hosts
fig_5a <- ggplot(df_m1, aes(x = scenario, y = avg, 
                            group = type, color = type)) +   
  geom_quasirandom(width = 0.1, dodge.width = 0.85, 
                   alpha = 0.1, size = 0.15, color = "grey80") + 
  geom_boxplot(aes(group = interaction(scenario, type)), 
               width = 0.75, position = position_dodge(0.85), 
               alpha = 0, outlier.alpha = 0, size = 0.4) + 
  facet_grid(~ roaming) +
  scale_y_continuous(limits = c(0, 175), breaks = seq(0, 150, by = 50)) +
  scale_color_manual(values = c("#4aa9a0", "#3a2cd7"), name = "Group:") + 
  labs(x = "Landscape scenario", y = "\nVariance in\nhost density", 
       tag = "(a)", title = NULL, subtitle = NULL) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 12.5))

## b) visits by ranging males
fig_5b <- ggplot(df_m2, aes(x = scenario, y = med, group = type, color = type)) +   
  geom_quasirandom(width = 0.1, dodge.width = 0.85, 
                   alpha = 0.25, size = 0.25, color = "grey80") + 
  geom_boxplot(aes(group = interaction(scenario, type)), 
               width = 0.75, position = position_dodge(0.85), 
               alpha = 0, outlier.alpha = 0, size = 0.4) + 
  facet_grid(~ roaming) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 3)) +
  scale_color_manual(values = c("#7aa6e5", "#1a3357"), name = "Period:") + 
  labs(x = "Landscape scenario", y = "\nVisited patches\nper week", 
       tag = "(b)", title = NULL, subtitle = NULL) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank())

## c) contact rates per quality class
fig_5c <- ggplot(df_m3, aes(x = scenario, y = med, group = type, color = type)) +   
  geom_quasirandom(width = 0.05, dodge.width = 0.95, 
                   alpha = 0.1, size = 0.1, color = "grey80") + 
  geom_path(data = hom, aes(x = scenario, y = mean, group = roaming), 
            color = "grey65", linetype = "dashed", size = 0.5) +
  geom_boxplot(aes(group = interaction(scenario, type)), width = 0.95, 
               position = position_dodge2(1, preserve = "single"), 
               alpha = 0, outlier.alpha = 0, size = 0.4) + 
  facet_grid(~ roaming) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25)) +
  scale_color_manual(values = c("#8cd085", "#4cad42", "#26642b"), 
                     name = "Habitat quality:") + 
  labs(x = "Landscape scenario", y = "\nContact rates of\nranging males", 
       tag = "(c)", title = NULL, subtitle = NULL) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank())

## d) simultaneously affected patches
fig_5d <- ggplot(df_m4, aes(x = scenario, y = max)) +   
  geom_quasirandom(width = 0.2, dodge.width = 0.175, 
                   alpha = 0.25, size = 0.25, color = "grey65") + 
  geom_boxplot(aes(group = scenario), width = 0.7,
               alpha = 0, outlier.alpha = 0, size = 0.4, color = "#cc5d00") + 
  facet_grid(~ roaming) +
  labs(x = "Landscape scenario", 
       y = "\nMaximum number of\nsimultaneously affected patches", 
       tag = "(d)", title = NULL, subtitle = NULL) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank())

## e) affected patches
fig_5e <- ggplot(df_m5, aes(x = scenario, y = max, group = scenario)) +   
  geom_quasirandom(width = 0.2, dodge.width = 0.175, 
                   alpha = 0.25, size = 0.25, color = "grey65") + 
  geom_boxplot(width = 0.7,
               alpha = 0, outlier.alpha = 0, size = 0.4, color = "#b22222") + 
  facet_grid(~ roaming) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  labs(x = "Landscape scenario", y = "\nProportion of\naffected patches", 
       tag = "(e)", title = NULL, subtitle = NULL) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank())

## f) trans within + between
fig_5f <- ggplot(df_m6, aes(x = scenario, y = avg, 
                            group = type, color = type)) +   
  geom_quasirandom(width = 0.1, dodge.width = 0.85, 
                   alpha = 0.1, size = 0.15, color = "grey80") + 
  geom_boxplot(aes(group = interaction(scenario, type)), 
               width = 0.75, position = position_dodge(0.85), 
               alpha = 0, outlier.alpha = 0, size = 0.4) + 
  facet_grid(~ roaming) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  scale_color_manual(values = c("#580080", "#d48d09"), 
                     name = "Transmission:") + 
  labs(x = "Landscape scenario", y = "\nProportion of\nsuccessful transmissions", 
       tag = "(f)", title = NULL, subtitle = NULL) +
  theme(axis.text.x = element_text(size = 9.5, angle = 33, 
                                   hjust = 1, vjust = 1),
        axis.title.x = element_text(size = 11),
        strip.text = element_blank(),
        strip.background = element_blank())
 
((fig_5 <- fig_5a + fig_5b + fig_5c + fig_5d + fig_5e + fig_5f) *
    theme(axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 9.5),
          plot.margin = margin(1, 1, 1, 1),
          plot.tag = element_text(size = 12),
          legend.title = element_text(size = 10, face = "plain"),
          legend.text = element_text(size = 9)) +
    plot_layout(ncol = 1))

## final panel
if (save == T) { 
  ggsave("./plots/final/Fig5_multi_box.png", width = 9, height = 11.5, dpi = 750) 
  ggsave("./plots/final/Fig5b_multi_box.pdf", width = 9, height = 11.5, device = cairo_pdf) 
}
```

### Fig. 4 v2: Median + Line

```{r fig-5-multi-median, fig.width = 9, fig.height = 11.5, eval = F, include = F}
## a) density of hosts
fig_5b_a <- df_m1  %>% 
  group_by(roaming, scenario, type) %>% 
  summarize(avg = mean(avg)) %>% 
  ggplot(aes(x = scenario, y = avg, group = type, color = type, shape = type)) +   
    geom_path(position = position_dodge(width = 0.55), size = 1, alpha = 0.4) +
    geom_point(position = position_dodge(width = 0.55), size = 2.5) +
    facet_grid(~ roaming) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 50)) +
    scale_color_manual(values = c("#69bdb5", "#4e41db"), name = "Group:") + 
    scale_shape_manual(values = c(19, 17), name = "Group:") +
    labs(x = "Landscape scenario", y = "\nVariance in\nhost density", 
         tag = "(a)", title = NULL, subtitle = NULL) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_text(size = 12))

## b) visits by ranging males
fig_5b_b <- df_m2 %>% 
  summarize(avg = mean(med, na.rm = T)) %>% 
  ggplot(aes(x = scenario, y = avg, group = type, color = type, shape = type)) +   
    geom_path(position = position_dodge(width = 0.55), size = 1, alpha = 0.4) +
    geom_point(position = position_dodge(width = 0.55), size = 2.5) +
    facet_grid(~ roaming) +
    scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 3)) +
    scale_color_manual(values = c("#7aa6e5", "#1a3357"), name = "Period:") + 
    scale_shape_manual(values = c(19, 17), guide = F) + 
    labs(x = "Landscape scenario", y = "\nVisited patches\nper week", 
         tag = "(b)", title = NULL, subtitle = NULL) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_blank(),
          strip.background = element_blank())

## c) contact rates per quality class
fig_5b_c <- df_m3  %>% 
  group_by(roaming, scenario, type) %>% 
  summarize(avg = mean(med)) %>% 
  ggplot(aes(x = scenario, y = avg)) +   
    geom_path(data = hom, aes(x = scenario, y = mean, group = roaming), 
              color = "grey60", linetype = "dashed", size = 0.5) +
    geom_path(aes(group = type, color = type),
              position = position_dodge(width = 0.55), size = 1, alpha = 0.4) +
    geom_point(aes(group = type, color = type, shape = type),
               position = position_dodge(width = 0.55), size = 2.5) +
    facet_grid(~ roaming) +
    scale_y_continuous(limits = c(0, 58), breaks = seq(0, 50, by = 25)) +
    scale_color_manual(values = c("#8cd085", "#3c8834", "#1c3f18"),
                       name = "Habitat quality:") + 
    scale_shape_manual(values = c(19, 17, 15), 
                       name = "Habitat quality:") + 
    labs(x = "Landscape scenario", 
         y = "\nContact rates of\nmoving hosts", 
         tag = "(c)", title = NULL, subtitle = NULL) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_blank(),
          strip.background = element_blank())

## d) infected patches
fig_5b_d <- df_m4 %>% 
  group_by(roaming, scenario) %>% 
  summarize(avg = mean(max, na.rm = T)) %>% 
  ggplot(aes(x = scenario, y = avg, 
             group = roaming)) +   
    geom_path(position = position_dodge(width = 0.55), 
              size = 1, alpha = 0.4, color = "#cc5d00") +
    geom_point(position = position_dodge(width = 0.55), size = 2.5) +
    facet_grid(~ roaming) +
    labs(x = "Landscape scenario", 
         y = "\nMaximum number of\nsimultaneously affected patches", 
         tag = "(d)", title = NULL, subtitle = NULL) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_blank(),
          strip.background = element_blank())

## e) affected patches
fig_5b_e <- df_m5 %>% 
  group_by(roaming, scenario) %>% 
  summarize(avg = mean(max, na.rm = T)) %>% 
  ggplot(aes(x = scenario, y = avg, group = roaming)) +   
    geom_path(position = position_dodge(width = 0.55), 
              size = 1, alpha = 0.4, color = "#a01e1e") +
    geom_point(position = position_dodge(width = 0.55), size = 2.5, color = "#a01e1e") +
    facet_grid(~ roaming) +
    scale_y_continuous(limits = c(0, 1), 
                       breaks = seq(0, 1, by = 0.25)) +
    labs(x = "Landscape scenario", 
         y = "\nProportion of\naffected patches", 
         tag = "(e)", title = NULL, subtitle = NULL) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_blank(),
          strip.background = element_blank())

## f) trans within + between
fig_5b_f <- df_m6 %>% 
  group_by(roaming, scenario, type) %>% 
  summarize(avg = median(avg, na.rm = T)) %>% 
  ggplot(aes(x = scenario, y = avg, 
             group = type, color = type, shape = type)) +   
    geom_path(position = position_dodge(width = 0.55), size = 1, alpha = 0.4) +
    geom_point(position = position_dodge(width = 0.55), size = 2.5) +
    facet_grid(~ roaming) +
    scale_y_continuous(limits = c(-0.1, 1.1), 
                       breaks = seq(0, 1, by = 0.25)) +
    scale_color_manual(values = c("#d48d09", "#580080", "#8a8a8a"), 
                       name = "Transmission:") + 
    scale_shape_manual(values = c(19, 17, 15), 
                       name = "Transmission:") + 
    labs(x = "Landscape scenario", 
         y = "\nProportion of\ninfections", 
         tag = "(f)", title = NULL, subtitle = NULL) +
    theme(axis.text.x = element_text(size = 9.5, angle = 33, 
                                     hjust = 1, vjust = 1),
          axis.title.x = element_text(size = 11),
          strip.text = element_blank(),
          strip.background = element_blank())

## final panel
(fig_5_b <- (fig_5b_a + fig_5b_b + fig_5b_c + fig_5b_d + fig_5b_e + fig_5b_f) *
    theme(axis.text.y = element_text(size = 9.5),
          axis.title.y = element_text(size = 11),
          plot.margin = margin(1, 1, 1, 1),
          plot.tag = element_text(size = 13),
          legend.title = element_text(size = 10, face = "plain"),
          legend.text = element_text(size = 9)) +
    plot_layout(ncol = 1))

if (save == T) { 
  ggsave("./plots/final/Fig5_v2_multi_med.png", width = 9, height = 11.5, dpi = dpi) 
  ggsave("./plots/final/Fig5_v2_multi_med.pdf", width = 9, height = 11.5, device = cairo_pdf) 
}
```


## Fig. 6: Probabilities of invasion success and persistence per case fatality risk

```{r fig-6-pers-cfr-tile, fig.width = 8, fig.height = 6.3}
df_6 <- df_infmean_plots %>% 
  mutate(roaming = factor(roaming, labels = lab_roam_n)) %>% 
  group_by(roaming, scenario, cfr) %>%
  summarize(
    pers_mean = mean(pers),
    inv_mean = mean(inv)
  ) %>% 
  ungroup()

fig_6a <- ggplot(df_6, aes(cfr, fct_rev(scenario))) +
  geom_tile(aes(fill = pers_mean), color = "white", size = 0.3) +
  facet_wrap(~ roaming) + 
  scale_fill_viridis(option = "viridis", name = "Probability", limits = c(0, 0.48),
                     breaks = seq(0, 0.4, by = 0.1)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Case fatality ratio", y = NULL, tag = "(a)", 
       title = "Long-term persistence (outbreak duration > 10 years)")

fig_6b <- ggplot(df_6, aes(cfr, fct_rev(scenario))) +
  geom_tile(aes(fill = inv_mean), color = "white", size = 0.3) +
  facet_wrap(~ roaming) + 
  scale_fill_viridis(option = "cividis", name = "Probability", limits = c(0, 1)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Case fatality ratio", y = NULL, tag = "(b)", 
       title = "Short-term persistence (outbreak duration > 2.5 years)")

(fig_6 <- (fig_6a + fig_6b) *
  theme(strip.text = element_text(size = 11),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.ticks.y = element_blank(),
        plot.title = element_text(size = 14),
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10, color = "grey40")) + 
  plot_layout(ncol = 1))

if (save == T) { 
  ggsave("./plots/final/Fig6_pers_cfr_v2.png", width = 8, height = 6.3, dpi = dpi) 
  ggsave("./plots/final/Fig6_pers_cfr_v2.pdf", width = 8, height = 6.3, device = cairo_pdf)
}
```


# Figures Supplement

## Fig. S3: Movement weights

```{r fig-s3-move-weights, fig.width = 12, fig.height = 6.15}
## CRW
x <- seq(-1, 1, length = 100)
d <- data.frame(
  x = x,
  y = -sqrt(1 - x^2),
  height = 2*sqrt(1 - x^2)
)

p_CRW <- ggplot(d, aes(x, y, height = height, group = 1, fill = x)) +
  geom_ridgeline_gradient(color = NA) +
  scale_fill_viridis_c(guide = "none", option = "cividis") +
  coord_flip() +
  labs(x = NULL, y = "", 
       title = "(a) Correlated random walk\n\n") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank())

## habitat-dependent
qual <- seq(0, 9, by = 0.01)
x <- seq(0, 2, by = 0.01)
HD <- expand.grid(qual, x)
colnames(HD) <- c("qual", "x")

titleH1 <- "\nHabitat quality B"
titleH2 <- ""

p_HD <- ggplot(HD, aes(x, qual)) +
  geom_tile(aes(fill = qual)) +
  scale_fill_viridis_c(option = "cividis") +
  scale_x_continuous(breaks = c(-5, 5), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 9, by = 1),expand = c(0, 0)) +
  labs(x = titleH2, y = titleH1, 
       title = "(b) Habitat-dependent\n    movement\n") +
  theme(legend.position = "none",
        axis.ticks = element_line(colour = "black", size = 0.25),
        panel.border = element_blank())

## density dependent
dens <- seq(0, 80, by = 0.1)
K <- seq(0, 40, by = 0.1)
DD <- expand.grid(dens, K)
colnames(DD) <- c("dens", "K")

titleD1 <- "\nNumber of conspecifics"
titleD2 <- "Carrying capacity K"

p_DD <- ggplot(DD, aes(dens, K)) +
  geom_tile(aes(fill = K - dens)) +
  scale_fill_viridis_c(option = "cividis") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 80, by = 10), expand = c(0, 0)) +
  labs(x = titleD2, y = titleD1,
       title = "(c) Competition-driven\n    movement\n") +
  theme(legend.position = "none",
        axis.ticks = element_line(colour = "black", size = 0.25),
        panel.border = element_blank())

## full plot
(fig_s3 <- (p_CRW + p_HD + p_DD) * theme(plot.title = element_text(size = 18)) + 
    plot_layout(ncol = 3, widths = c(3.5, 1, 3)))

if (save == T) { 
  ggsave("./plots/final/FigS3_moverules.png", width = 12, height = 6.15, dpi = dpi)
  ggsave("./plots/final/FigS3_moverules.pdf", width = 12, height = 6.15, device = cairo_pdf)
}
```


## Fig. S7: Persistence probabilities per case fatality ratio and persistence threshold

```{r fig-s7-facet-cf, fig.width = 7, fig.height = 8}
df_s7 <- df_infmean_plots %>% 
  mutate(
    pers_5 = if_else(duration_weeks > 5*52, 1, 0),
    pers_7 = if_else(duration_weeks > 7.5*52, 1, 0),
    roaming = factor(roaming, labels = lab_roam_n)
  ) %>% 
  group_by(roaming, scenario, cfr) %>%
  summarize(
    p2 = mean(inv),
    p5 = mean(pers_5),
    p7 = mean(pers_7),
    p10 = mean(pers)
  ) %>% 
  ungroup() %>% 
  gather(threshold, prob, -roaming, -scenario, -cfr) %>% 
  mutate(
    threshold = case_when(
      threshold == "p2" ~ "> 2.5 years", 
      threshold == "p5" ~ "> 5 years",
      threshold == "p7" ~ "> 7.5 years",
      threshold == "p10" ~ "> 10 years"),
    threshold = factor(threshold, 
                       levels = c("> 2.5 years", "> 5 years", 
                                  "> 7.5 years", "> 10 years")))

(fig_s7 <- ggplot(df_s7, aes(cfr, fct_rev(scenario))) +
  geom_tile(aes(fill = prob), color = "white", size = 0.33) +
  facet_grid(threshold ~ roaming) + 
  scale_fill_viridis(name = "Persistence probability", limits = c(0, 1),
                     guide = guide_colourbar(direction = "horizontal",
                                            barheight = unit(4, units = "mm"), 
                                            barwidth = unit(75, units = "mm"),
                                            draw.ulim = FALSE, title.position = 'top',
                                            title.hjust = 0.5, label.hjust = 0.5)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Case fatality ratio", y = NULL) + 
  theme(strip.text = element_text(size = 11),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 11),
        axis.ticks.y = element_blank(),
        plot.title = element_text(size = 14),
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10, color = "grey40"),
        legend.position = "bottom"))

if (save == T) { 
  ggsave("./plots/final/FigS7_cfr_persall_v2.png", width = 7, height = 8, dpi = dpi) 
  ggsave("./plots/final/FigS7_cfr_persall_v2.pdf", width = 7, height = 8, device = cairo_pdf) 
}
```


## Fig. S8: Outbreak duration (time to pathogen extinction)

```{r fig-s8-time2extinction, fig.width = 12, fig.height = 7.1}
n <- df_infmean_plots %>%   
  group_by(roaming, cfr, scenario) %>%
  count() %>% 
  ungroup() %>% 
  summarize(n = unique(n)) %>% 
  pull()

df_s8 <- df_infmean_plots %>% 
  filter(cfr == cfr_filter) %>% 
  mutate(duration_quarter = if_else(duration_quarter > 40, 50, 
                                    floor(duration_quarter))) %>% 
  group_by(roaming, scenario, duration_quarter) %>% 
  summarize(prop = n() / n)

(fig_s8 <- ggplot(filter(df_s8, duration_quarter != 50), 
                  aes(duration_quarter / 4, prop)) + 
  geom_bar(aes(group = duration_quarter, fill = scenario), 
           stat = "identity", width = 0.2) +
  geom_bar(data = filter(df_s8, duration_quarter == 50), 
           aes(group = duration_quarter, fill = scenario), 
           stat = "identity", width = 0.2) +
  geom_vline(xintercept = 11.25, linetype = "dotted") +
  facet_grid(scenario ~ roaming) +
  scale_fill_viridis_d(option = "inferno", begin = 0.85, end = 0.15, guide = F) +
  scale_x_continuous(breaks = seq(0, 12.5, by = 2.5),
                     labels = c(0, 2.5, 5, 7.5, 10, "> 10 years")) +
  scale_y_continuous(breaks = seq(0, 0.4, 0.2)) +
  expand_limits(x = c(0, 13), y = c(0, 0.41)) +
  theme(panel.grid.major.y = element_line(color = "grey90", size = 0.4),
        panel.grid.minor.y = element_line(color = "grey90", size = 0.4)) +
  labs(x = "Outbreak duration (time to pathogen extinction)", 
       y = "Proportion of simulation runs"))

if (save == T) { 
  ggsave("./plots/final/FigS8_exttime.png", width = 12, height = 7.1, dpi = dpi) 
  ggsave("./plots/final/FigS8_exttime.pdf", width = 12, height = 7.1, device = cairo_pdf) 
}
```


## Fig. S9: Temporal dynamics of transmission rates by group-living and ranging individuals

```{r fig-s9-trans-males+group, fig.width = 9, fig.heigth = 6.7}
(fig_s9 <- df_trans %>% 
  filter(trans %in% c("g", "m")) %>%
  ggplot(aes(quarter / 4, avg)) +
    geom_vline(aes(xintercept = 2.625), 
               linetype = "dashed", color = "grey60", size = 0.3) +
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = type), 
                alpha = 0.25) + 
    geom_line(aes(color = type), size = 0.5) +
    geom_rug(data = df_ext, aes(x = week_ext / 52), 
             sides = "b", color = "grey60", alpha = 0.5, size = 0.3) +
    facet_grid(scenario ~ roaming) +
    scale_x_continuous(breaks = seq(0, 10, by = 2.5), 
                       labels = seq(0, 10, by = 2.5)) +
    scale_y_continuous(limits = c(-0.001, 0.25), 
                       breaks = seq(0, 0.3, by = 0.1)) +
    scale_color_manual(values = c("#7cdc9e", "#766ce3"), name = NULL) +
    scale_fill_manual(values  = c("#7cdc9e", "#766ce3"), name = NULL) +
    labs(x = "Outbreak duration (years)", 
         y = "Weekly transmissions per individual") +
    theme(legend.position = "bottom",
          strip.text.x = element_text(size = 12),
          strip.text.y = element_text(size = 11.5),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          axis.text = element_text(size = 10),
          legend.text = element_text(size = 10.5)))

if (save == T) { 
  ggsave("./plots/final/FigS9_trans_g+m.png", width = 9, height = 6.7, dpi = dpi) 
  ggsave("./plots/final/FigS9_trans_g+m.pdf", width = 9, height = 6.7, device = cairo_pdf)
}
```


## Fig. S10: Density pre- and post release

```{r}
df_all %>% 
  filter(cfr == 0.5) %>% 
  mutate(
    roaming = factor(roaming, labels = lab_roam),
    scenario = factor(scenario, 
                      levels = c("hom", "patch_l", "patch_s", "rand"), 
                      labels = lab_scen),
    period = ifelse(week < week_release, "Pre-Outbreak", "Outbreak"),
    period = factor(period, levels = c("Pre-Outbreak", "Outbreak"))
  ) %>% 
  group_by(roaming, scenario, run, period) %>% 
  summarize(males = median(dens_var_roam, na.rm = T)) %>% 
  ggplot(aes(x = scenario, y = males, 
             group = period, color = period)) +   
    geom_quasirandom(width = 0.1, dodge.width = 0.9, 
                     alpha = 0.3, size = 0.7, color = "grey75") + 
    geom_boxplot(aes(group = interaction(scenario, period)), 
                 width = 0.8, position = position_dodge(0.9), 
                 alpha = 0, outlier.alpha = 0, size = 0.4) + 
    facet_grid(~ roaming) +
    scale_y_log10()+
    scale_color_manual(values = c("#3fa78e", "#205455"), 
                       name = "Period:") + 
    labs(x = "Landscape scenario", 
         y = "Variance in density of moving hosts") +
    theme(axis.text.x = element_text(size = 12, angle = 33, 
                                     hjust = 1, vjust = 1),
          legend.title = element_text(size = 11.5, face = "plain"),
          legend.text = element_text(size = 10),
          legend.position = "bottom",
          strip.text = element_text(size = 12))

if (save == T) { 
  ggsave("./plots/final/FigS10_density_pre+post.png", width = 9, height = 5, dpi = 750)
  ggsave("./plots/final/FigS10_density_pre+post.pdf", width = 9, height = 5, device = cairo_pdf) 
}
```


## Fig. S11: Infected individuals over time

```{r fig-s11-inf-ind-time, fig.width = 12, fig.height = 9}
df_s11 <- df_inf_plots %>% 
  filter(cfr == cfr_filter) %>% 
  group_by(scenario, roaming) %>% 
  mutate(pers = ifelse(duration_weeks > 520, ">10", "<10")) %>% 
  group_by(scenario, roaming, week_pathogen) %>% 
  mutate(
    meanInf  = mean(ind_inf),
    duration_weeks = if_else(duration_weeks == week_pathogen, duration_weeks, NA_real_)
  )

labs <- c("\u2264 10 years", "> 10 years")

fig_s11 <- ggplot() +   
  geom_line(data = filter(df_s11, pers == ">10"), aes(week_pathogen, ind_inf, group = run, 
            color = ">10"), size = 0.3, alpha = 0.4) +
  geom_line(data = filter(df_s11, pers == "<10"), aes(week_pathogen, ind_inf, group = run, 
            color = "<10"), size = 0.3, alpha = 0.4) +
  geom_line(data = df_s11, aes(week_pathogen, meanInf), color = "black", size = 0.35) +
  geom_vline(xintercept = c(130, 260, 390, 520), linetype = "dashed", size = 0.2, color = "grey40") +
  geom_rug(data = filter(df_s11, pers == "<10"), aes(duration_weeks), 
           sides = "b", alpha = 0.4, size = 0.3) +
  facet_grid(scenario ~ roaming) +
  scale_x_continuous(breaks = seq(0, 520, by = 52), limits = c(0, 520), 
                       labels = 0:10) + 
  scale_y_continuous(labels = scales::comma_format()) +
  scale_color_manual(values = c("<10" = "grey20", ">10" = "tan1"), 
                     name = "Outbreak duration:", labels = labs) +
  guides(color = guide_legend(override.aes = list(size = 0.6, alpha = 1))) +
  theme(legend.position = c(0.235, 0.92), 
        legend.box = "horizontal",
        legend.title = element_text(size = 10, hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.key = element_rect(fill = NA, color = NA),
        legend.key.width = unit(2.2, "lines"),
        legend.background = element_rect(fill = "white", color = "grey40", size = 0.3),
        plot.background = element_rect(fill = NA, color = NA),
        panel.grid.major.y = element_line(color = "grey85", size = 0.3),
        panel.grid.minor.y = element_line(color = "grey85", size = 0.3)) +
  labs(x = "Years since pathogen release", y = "Infected individuals")

if (save == T) { 
  ggsave("./plots/final/FigS11_dynamics_inf.png", width = 12, height = 9, dpi = 750)
  ggsave("./plots/final/FigS11_dynamics_inf.pdf", width = 12, height = 9, device = cairo_pdf) 
}
```


## Fig. S12: Affected habitat cells over time

```{r fig-s12-inf-cells-time, fig.width = 11.5, fig.height = 8}
df_quarter %>% 
  dplyr::select(run, roaming, scenario, week_pathogen, cell_infectious, patches_0) %>% 
  mutate(roaming = factor(roaming, levels = lab_roam_n, labels = lab_roam)) %>% 
  group_by(roaming, scenario, week_pathogen) %>% 
  mutate(
    prop_cell_infectious = cell_infectious / (1250 - patches_0),
    avg = sum(prop_cell_infectious) / n()
  ) %>% 
  ggplot(aes(week_pathogen, prop_cell_infectious)) + 
    #geom_vline(xintercept = 78, color = "grey40", size = 0.3, linetype = "dashed") +  ## for Rev. 2
    #geom_vline(xintercept = 130, color = "grey40", size = 0.3) +  ## for Rev. 2
    geom_line(aes(group = run), color = "grey30", size = 0.4, alpha = 0.2) + 
    geom_line(aes(week_pathogen, avg), color = "#cc5d00", size = 0.7) +
    facet_grid(scenario ~ roaming) +
    scale_x_continuous(breaks = seq(0, 520, by = 52), limits = c(0, 520), 
                       labels = 0:10, expand = c(0.01, 0.01)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
    theme(panel.grid.major.y = element_line(color = "grey70", size = 0.3, linetype = "dotted"),
          axis.text = element_text(size = 10)) +
    labs(x = "Years since pathogen release", y = "Number of affected habitat patches")

if (save == T) { 
  ggsave("./plots/final/FigS12_affected_cells.png", width = 11.5, height = 8, dpi = 750)
  ggsave("./plots/final/FigS12_affected_cells.pdf", width = 11.5, height = 8, device = cairo_pdf) 
}
```


## Fig. S13: Cumulative number of affect chabitat cells

```{r fig-s13-cum-inf-cells-time, fig.width = 11.5, fig.height = 8}
df_quarter %>% 
  dplyr::select(run, roaming, scenario, week_pathogen, cell_infected, patches_0) %>% 
  mutate(roaming = factor(roaming, levels = lab_roam_n, labels = lab_roam)) %>% 
  group_by(roaming, scenario, week_pathogen) %>% 
  mutate(
    prop_cell_infected = cell_infected / (1250 - patches_0),
    avg = sum(prop_cell_infected) / n()
  ) %>% 
  ggplot(aes(week_pathogen, prop_cell_infected)) + 
    #geom_vline(xintercept = 78, color = "grey40", size = 0.3, linetype = "dashed") +  ## for Rev. 2
    geom_line(aes(group = run), color = "grey30", size = 0.4, alpha = 0.2) + 
    geom_line(aes(week_pathogen, avg), color = "#b22222", size = 0.7) +
    facet_grid(scenario ~ roaming) +
    scale_x_continuous(breaks = seq(0, 520, by = 52), limits = c(0, 520), 
                       labels = 0:10, expand = c(0.01, 0.01)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
    theme(panel.grid.major.y = element_line(color = "grey70", size = 0.3, linetype = "dotted"),
          axis.text = element_text(size = 10)) +
    labs(x = "Years since pathogen release", y = "Cumulative number of affected habitat patches")

if (save == T) { 
  ggsave("./plots/final/FigS13_affected_cells_cum.png", width = 11.5, height = 8, dpi = 750)
  ggsave("./plots/final/FigS13_affected_cells_cum.pdf", width = 11.5, height = 8, device = cairo_pdf) 
}
```

***

# Version info

```{r version}
version
sessionInfo()
```
